\chapter{Introduction}
\label{chap:prelim}

\section{Sequence and network data}

	\subsection{Genome, transcriptome, proteome}

		\subsubsection{Gene expression}

		\paragraph{Measuring gene expression levels}

		\paragraph{Differential analysis}

			\begin{itemize}
				\item better understanding of cellular processes
				\item biomarkers discovery
			\end{itemize}

	\subsection{Biological networks}
		Biological networks are abstract representations of biological entities interconnected by some criteria.
		They can represent for example the relationships between species inside an ecosystem, or interconnections between cell types in any multicellular organism.

		In this work, we are mostly interested in networks at the biomolecular level.
		Many such networks exists, to name a few:
		\begin{itemize}
			\item \emph{Metabolic networks} represent biochemical reactions between substrates, enzymes and metabolites, and cluster them into pathways,
			\item \emph{Gene co-expression networks} represent the similarity of expression between genes in some biological setup, by interconnecting pairs of genes simultaneously expressed,
			\item \emph{Gene regulatory network} represent the indirect regulatory actions of genes, from proteins and transcription factors to gene expression levels,
			\item \emph{Protein-protein interaction networks} represent interactions between two proteins, usually of the same species.
		\end{itemize}

		Unfortunately, there exists some ambiguity in the uses of the term biological networks, and biological networks can be seen as a combination of observed and inferred facts.
		Usually the context can help understand the variations in meaning, but let us explicit the two most understood meanings here.
		It is not uncommon for biologists to treat biological networks as observed biological facts.
		In these uses, the network \emph{is the knowledge}, usually some well observed biological interaction; e.g. a known pathway that connect chemical reactants and products through enzymes.
		On the other hand, mostly in computational contexts, the networks are \emph{abstract representation of the knowledge}.
		There, nodes represent entities and edges represent some form of deduced connections; e.g. a gene co-expression network, which is statistically constructed from control and condition genes expression profiles.

		Let us stress the importance of biological networks in modern biology.
		They structure our understanding of biological systems and both allow a comprehension of biological processes at the systems level, and permit automated processing of the knowledge that they represent.
% XXX TODO		As automated processing enabling tools, they can serve as both knowledge bases for local decisions and as global networks that can serve as XXX substrate (de quoi parles-tu? -- Macha) XXX for integrated analysis.

		XXX.

		The most fitting abstraction for those biological networks are discrete mathematics' graphs (XXX to be formally defined in ...XXX).

		Protein-protein interactions (PPI) networks play an important role in this work, and we will present them in more detail.

		\subsubsection{Protein-Protein Interactions}


%	\subsubsection{??? Phage display ???}
%	\subsubsection{??? Mass spectrometry ???}

\section{Elements of graph theory}

	\subsection{Graphs}
		Let us recall some basic material related to graphs.
		A graph $G = (V,E)$ consists of a set of vertices $V$ and a set of edges (unordered pairs of vertices) $E$.
		%To shorten the exposition, we shall usually abbreviate $|V|$ and $|E|$ to $n$ and $m$, respectively.
		We say that $G$ is node-weighted if a function $w\colon V \to \mR$ is provided.
		%A graph is a \textit{tree} if it is both connected -- there exists a path between any pair of vertices -- and acyclic -- it does contain a closed path in which the first and the last vertices are the same. 
		Given a graph $G = (V, E)$, its subgraph $G' = (V', E')$ is said to be \emph{induced} if $G'$ has exactly the edges that appear in $G$ over the vertex set $V' \in V$, that is $E' = \Set{(x, y) \in E}{x,y \in V'}$.
		We  denote the graph \emph{induced} by the node set $V'$ in $G$ by $G\left[V'\right]$.

		\paragraph{Minimum cut}

\section{Combinatorial optimization}
	\paragraph{Dynamic programming}
	\paragraph{Decision trees, Branch and bound, Branch and cut}
	\paragraph{Linear programming, Mixed integer linear programming}

	\subsection{Complexity}
		\paragraph{APX-hardness}
			\label{par:m3sat}
			One well studied APX-hard problem is the \msat{} problem and is defined by \textcite{papadimitriou1991optimization} as follows.
			Given a collection $C_q = \{c_1, \ldots c_q\}$ of $q$ clauses where each clause consists of a set of three literals over a finite set of $n$ boolean variables $V_n = \{x_1, \ldots x_n\}$ and every literal occurs in at most $B$ clauses, is there a truth assignment of $V_n$ satisfying the largest number of clauses of $C_q$?

		\paragraph{Pseudo-polynomial time}

%	\subsection{String}
%		\paragraph{Suffix trees and array}
